---
title: "writing_functions"
output: html_document
date: "2025-10-23"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(readxl)
set.seed(1)

knitr::opts_chunk$set(
  echo=TRUE,
  warning = FALSE
)
```

every one loves z scores

```{r}
x_vec=rnorm(20, 10,2)
(x_vec - mean(x_vec))/sd(x_vec)

```
writing a function to compute Z score.

```{r}
z_score = function(x) {
  
  if(!is.numeric(x)) {
    stop("The input x should be numeric")
  }
  
  if(length(x)<5){
    
    stop("Only compute z scores when the input has 5 or more numbers")
  }
  
  z=(x - mean(x))/sd(x)
  
  z
  
}
```

let's try our function

```{r}
z_score(x=x_vec)
num_vec=rnorm(123,14,0.4)
z_score(x=num_vec)
```

let's break our function
```{r error=TRUE}
z_score(3)
z_score("ss")
```
## Let's compute some stuff

let's compute and return the mean and sd of a numeric vector

```{r}
mean_sd=function(x){
  
   if(!is.numeric(x)) {
    stop("The input x should be numeric")
  }
  
  if(length(x)<5){
    stop("Only compute mean and standard deviation when the input has 5 or more numbers")
  }
  
  mean_x=mean(x, na.rm=TRUE)
  sd_x=sd(x, na.rm=TRUE)
  
  tibble(mean = mean_x, sd= sd_x)
  
  
}

mean_sd(x=x_vec)

```

## make up data

data simulation

```{r}
sim_df=
  tibble(
    x=rnorm(30,3,1)
  )

sim_df |> 
  summarize(
    mu_hat=mean(x),
    sigma_hat=sd(x)
  )
```
let's write a function to do simulation
 
Function simulates data from a normal and computes sample mean and sd.
```{r}

sim_md=function(n1=30, mu=3, sigma=2){
  
  sim_df=
    tibble(
      x=rnorm(n=n1,mean=mu,sd=sigma)
    )
  
  sim_df |> 
    summarize(
      mu_hat=mean(x),
      sigma_hat=sd(x)
    )
  }

```

let's run this function


```{r}
sim_md(n=30)

```

don't forget documentation

Import the LoTR data

```{r}

fellow_r <- read_excel("data/LotR_Words.xlsx", range = "B3:D6") |> 
  mutate(movie = "Fellow")


two <- read_excel("data/LotR_Words.xlsx", range = "F3:H6") |> 
  mutate(movie = "two")

return_k <- read_excel("data/LotR_Words.xlsx", range = "J3:L6") |> 
  mutate(movie = "Return")

  
```

```{r}

data_import=function(cell_range, movie_title){
  df=read_excel("data/LotR_Words.xlsx", range=cell_range) |> 
    mutate(movie=movie_title)
  df
}

fellowship= data_import("B3:D6", "Fellowship")
two=data_import("F3:H6", "Two tower")
```
```{r}


nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(2) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```
write an import function

```{r}
ns_import=function(html, table_num){

data = 
 html |> 
  html_table() |> 
  nth(table_num) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}

ns_import(nsduh_html, table_num=1)



```

